Demo: Counterpoint by Construction

### 2019-05-24 doc/farm/farm-abstract.tex
more tweaks

\section{Introduction}

Text of paper \ldots

%% Acknowledgments
\begin{acks}                            %% contents suppressed with 'anonymous'
\end{acks}

%% Bibliography
%\bibliography{bibfile}

>>>>

Western music of the common practice period tends to loosely
follow sets of rules, which were developed over time to ensure
the aethetic quality of the composition. 
Among various rules, those for harmony \citep{piston1987harmony}
and counterpoint \citep{fux1965study} (a technique for generating
multi-voice melodies) are continue to be taught to music students,
not only as a means to understand the music of that period, but
also as a foundation for modern art and popular music.

To help analyze and synthesize tonal music, it is worthwhile to
encode these rules into a programming language, and as shown by
recent studies, functional programming languages are particularly
suited to this task.
In the past decade, Haskell has been extensively used to encode
the rules of harmony \citep{fmmh, hihseufha, faamh, hafha, fghm}
as well as counterpoint \citep{szamozvancev2017well}.
Since Haskell is a statically typed language, these frameworks 
often come with a form of type safety property, namely
``well-typed music does not sound wrong''.
Unfortunately, it turns out that the type system of plain Haskell
is sometimes not powerful enough to guarantee this property.
To avoid generating music that ``sounds wrong'', previous studies
have incorporated various language extensions, such as GADTs 
\citep{cheney2002lightweight} and singleton types 
\citep{eisenberg2013dependently}, to enable dependently-typed 
programming.

In this demonstration of work in progress, we present Music Tools
\citep{MusicTools}, a library of small tools that can be combined
functionally to help analyze and synthesize music.
To allow simple and natural encoding of rules, we built the library
in Agda \citep{norellphd}, which is a functional language with full
dependent types.
As an application of the library, we demonstrate an implementation of 
species counterpoint, based on the rules given by \citet{fux1965study}.
Thanks to Agda's rich type system, we can ensure by construction that
well-typed counterpoint satisfies all the required rules.
We show how the type-based approach both aids human composition and
allows for computer-generated creation of correct counterpoint.
We also contrast our work to a recent study on generating
natural-sounding counterpoint by machine learning
\citep{CounterpointByConvolution}, which does not give us correctness
guarantees.
Finally, we discuss further applications of our library, including
handling of functional harmony.

%% Acknowledgments
\begin{acks}                            %% contents suppressed with 'anonymous'
 The authors would like to thank the participants of the Tokyo Agda 
 Implmentors' Meeting, especially Ulf Norell and Jesper Cockx,
 for many helpful suggestions that improved our Agda code.
\end{acks}

%% Bibliography
\bibliography{farm-abstract.bib}
